---
title: "Tau-defined subtypes by sex and metabolics"
author: "Brandon Hall"
output:
  prettydoc::html_pretty:
    theme: hpster
    highlight: github
    toc: true
    number_sections: true
---

# Loading libraries
```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2) # you know this legend
library(plyr) 
library(dplyr)
library(ggsignif) # for creating plots with sigificance added
library(nnet) # for creating multinomial logistic regression models
library(AICcmodavg) # for model comparison
library(tidyverse)
```

# Getting set up
## Loading data

This chunk loads the required CSVs and merges them into one dataframe
```{r data, echo=FALSE, message=TRUE, warning=FALSE}
setwd("~/Documents/BraskieLab/My projects/Tau subtypes x metabolics/")
df.subtypes <- read.csv("Subject tau subtype PET data v3.csv")
colnames(df.subtypes)[1] <- "Med_ID"
df.HABS <- read.csv("~/Documents/BraskieLab/HABS data/Data Release 3/HD 1  50+ Merged v2.csv")
df.all<- merge(df.subtypes, df.HABS, by = "Med_ID")
#write.csv(df.all, file = "~/Documents/BraskieLab/My projects/Tau subtypes x metabolics/subtype+HABS_merged.csv", row.names = FALSE)
```

### Change variable types

This chunk changes the variable types as needed for subsequent analyses
```{r var_types, echo=FALSE, message=FALSE, warning=FALSE}
df.all$CDX_Cog <- as.factor(df.all$CDX_Cog)
df.all$Age <- as.numeric(df.all$Age)
df.all$ID_Gender <- as.factor(df.all$ID_Gender)
df.all$CDX_Diabetes <- as.factor(df.all$CDX_Diabetes)
df.all$Diabetes <- as.factor(df.all$Diabetes)
df.all$CDX_Hypertension <- as.factor(df.all$CDX_Hypertension)
df.all$ID_Hispanic <- as.factor(df.all$ID_Hispanic)
df.all$ID_Race_Black <- as.factor(df.all$ID_Race_Black)
df.all$APOE4_Positivity <- as.factor(df.all$APOE4_Positivity)
df.all$simoa_plasma_Total_Tau <- as.numeric(df.all$simoa_plasma_Total_Tau)
df.all$Final_subtype <- factor(df.all$Final_subtype)
df.all$X01_AB_FBB_pos <- as.factor(df.all$X01_AB_FBB_pos)
df.all$APOE4_Positivity <- as.factor(df.all$APOE4_Positivity)
df.all$OM_BMI <- as.double(df.all$OM_BMI)
df.all$Subtype <- factor(df.all$Subtype, ordered = FALSE)
df.all$Age_log <- log(df.all$Age)
df.all$OM_BMI_log <- log(df.all$OM_BMI)
df.all$X01_L_hippocampus <- as.numeric(df.all$X01_L_hippocampus)
df.all$X01_R_hippocampus <- as.numeric(df.all$X01_R_hippocampus)
df.all$whole_hippo <- (df.all$X01_L_hippocampus + df.all$X01_R_hippocampus)/2
```

### Subset from MCI & AD

This chunk create subset dataframes from the prior, unified dataframe
```{r subsetting2}
# only MCI and AD subjects
df.mciad <- subset(df.all,df.all$CDX_Cog != "0")
df.mciad$OM_BMI <- as.double(df.mciad$OM_BMI)
df.mciad$CDX_Cog <- factor(df.mciad$CDX_Cog, ordered = FALSE)
df.females <- subset(df.mciad, df.mciad$ID_Gender == "1")
df.males <- subset(df.mciad, df.mciad$ID_Gender == "0")
df.submed <- subset(df.mciad, df.mciad$Final_subtype == "0")
df.typical <- subset(df.mciad, df.mciad$Final_subtype == "1")
df.lp <- subset(df.mciad, df.mciad$Final_subtype == "2")
df.hs <- subset(df.mciad, df.mciad$Final_subtype == "3")
df.MA <- subset(df.mciad, df.mciad$ID_Hispanic != "1")
df.NHW <- subset(df.mciad, df.mciad$ID_Hispanic == "1") 
df.black <- subset(df.mciad, df.mciad$ID_Race_Black == "1")
df.typical_f <- subset(df.typical, df.typical$ID_Gender=="1") # create typical female-only subset
df.typical_m <- subset(df.typical, df.typical$ID_Gender=="0") # create typical male-only subset
df.lp_f <- subset(df.lp, df.lp$ID_Gender=="1") # create lp female-only subset
df.lp_m <- subset(df.lp, df.lp$ID_Gender=="0") # create lp male-only subset
df.hs_f <- subset(df.hs, df.hs$ID_Gender=="1") # create hs female-only subset
df.hs_m <- subset(df.hs, df.hs$ID_Gender=="0") # create hs male-only subset
df.submed_f <- subset(df.submed, df.submed$ID_Gender=="1") # create submed female-only subset
df.submed_m <- subset(df.submed, df.submed$ID_Gender=="0") # create submed male-only subset

#write.csv(df.mciad, file = "~/Documents/BraskieLab/My projects/Tau subtypes x metabolics/MCI_AD.csv", row.names = FALSE)

df.test <- subset(df.mciad, select = c("Med_ID", "Subtype", "OM_BMI", "OM_BMI_log", "ID_Hispanic", "ID_Race_Black", "CDX_Diabetes", "Diabetes", "CDX_Hypertension", "Age", "Sex", "CDX_Cog", "Ratio.of.Hippocampal.SUVR.to.Neocortical.SUVR", "ID_Hispanic", "ID_Race_Black")) # create a subsetted data frame containing only what variables I need

df.test <- df.test[complete.cases(df.test) ,] # removed any subjects who have incomplete data
```

# Barcharts

## 1: Sex

This chunk provides the barcharts for visualizing the difference in sex between subtypes
```{r sex_plot}
# subtype x sex
ggplot(df.test, aes(x = Subtype, fill = Sex)) + 
  geom_bar(width = 0.4, position = "dodge") + 
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Sex by tau subtype", fill = "Sex", x = "Subtype") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
## 2: Diabetes

This chunk provides the barcharts for visualizing the difference in diabetic diagnoses between subtypes
```{r sex_diab_plot}

# subtype x diabetes
ggplot(df.test, aes(x = Subtype, fill = Diabetes)) + 
  geom_bar(width = 0.6, position = "dodge") + 
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Diabetic diagnosis by subtype", fill = "Diabetes", x = "Subtype") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

##3: hypertension

```{r hypertenseplot}

ggplot(df.mciad, aes(x = Subtype, fill = CDX_Hypertension)) + geom_bar(width = 0.4, position = "dodge") + labs(title = "hypertension diagnosis by subtype", fill = "hypertension diagnosis", x = "subtype") + ylim(0,100)
```
##4: ID_Black

```{r blackplot}

ggplot(df.test, aes(x = Subtype, fill = ID_Race_Black)) + 
  geom_bar(width = 0.4, position = "dodge") + 
  labs(title = "Black identity by subtype", fill = "Self-report identity: black", x = "subtype") + 
  ylim(0,100) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

##5: ID_Hispanic

```{r hispplot}

ggplot(df.test, aes(x = Subtype, fill = ID_Hispanic)) + 
  geom_bar(width = 0.4, position = "dodge") + 
  labs(title = "Hispanic identity by subtype", fill = "Self-report identity: Hispanic", x = "subtype") + 
  ylim(0,100) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
# Box/Violin city

## BMI
```{r}
ggplot(df.mciad, aes(x = Subtype, y = OM_BMI)) +
  geom_violin(aes(fill = Subtype), alpha = 0.5, draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(title = "BMI by subtype", y = "Body mass index") +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, vjust = 1)) + 
  geom_signif(comparisons = list(c("hippocampal sparing", "typical"), c("typical", "limbic predominant"), c("limbic predominant", "hippocampal sparing"), c("typical", "sub-median tau")), y_position = c(70,75, 80, 85))
```

## Age
```{r}
ggplot(df.mciad, aes(x = Subtype, y = Age)) + 
  geom_violin(aes(fill = Subtype), alpha = 0.5, draw_quantiles = c(0.25, 0.5, 0.75)) + 
  labs(title = "Age by subtype", y = "Age") + 
  theme(axis.text.x = element_text(size = 12,angle = 45, hjust = 1, vjust = 1)) + 
  geom_signif(comparisons = list(c("hippocampal sparing", "typical"), c("typical", "limbic predominant"), c("limbic predominant", "hippocampal sparing"), c("typical", "sub-median tau")), y_position = c(90,95,100,96))
```
## Age x Sex x Diabetes
```{r}
ggplot(df.test, aes(x = Subtype, y = Age, fill = Subtype)) + geom_boxplot(size = 0.75) + facet_grid(CDX_Diabetes ~ Sex, margins = FALSE) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
##BMI x Sex x Diabetes
```{r}
ggplot(df.test, aes(x = Subtype, y = OM_BMI, fill = Subtype)) + geom_boxplot(size = 0.75) + facet_grid(CDX_Diabetes ~ Sex, margins = FALSE) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

# T-tests
## Age

```{r t.test}
t.test(df.males$Age, df.females$Age) # total group t-mciad -- NS
t.test(df.typical$Age, df.submed$Age) # typical vs sub-med -- significant!
t.test(df.typical$Age, df.hs$Age) # typical vs hippocampal sparing -- NS
t.test(df.typical$Age, df.lp$Age) # typical vs limbic predominant -- NS
t.test(df.hs$Age, df.lp$Age) # hippocampal sparing vs limbic predomiant -- significant!

```
## BMI
```{r t.test2}
t.test(df.hs$OM_BMI, df.typical$OM_BMI)
t.test(df.lp$OM_BMI, df.typical$OM_BMI)
t.test(df.hs$OM_BMI, df.lp$OM_BMI)
t.test(df.submed$OM_BMI, df.typical$OM_BMI)
```

# Chi square test

``` {r chisq}
set.seed(1001)
chisq.test(df.mciad$ID_Hispanic, df.mciad$Subtype, correct = FALSE, p = rep(1/length(df.mciad$Subtype), length(df.mciad$Subtype)), simulate.p.value = TRUE, B = 1000)

chisq.test(df.mciad$ID_Race_Black, df.mciad$Subtype, correct = FALSE, p = rep(1/length(df.mciad$Subtype), length(df.mciad$Subtype)), simulate.p.value = TRUE, B = 1000)
```

# multinomial logistic regression

All models include covariates of age and cognitive status

## Typical reference
```{r multinom}
#library(nnet) # load the nnet library

#df.test <- subset(df.mciad, select = c("Med_ID", "Subtype", "OM_BMI", "OM_BMI_log", "ID_Hispanic", "ID_Race_Black", "CDX_Diabetes", "Diabetes", "CDX_Hypertension", "Age", "Sex", "CDX_Cog", "Ratio.of.Hippocampal.SUVR.to.Neocortical.SUVR", "ID_Hispanic", "ID_Race_Black")) # create a subsetted data frame containing only what variables I need

#df.test <- df.test[complete.cases(df.test) ,] # removed any subjects who have incomplete data

set.seed(0001)

df.test$Subtype <- relevel(df.test$Subtype, ref = "typical") # relevel the unordered factor such that "typical" is the reference level
m1 <- multinom(data = df.test, Subtype ~ CDX_Diabetes + CDX_Hypertension + OM_BMI_log + Age + Sex) # screate a multinomial logistic regression model to predict Subtype using CDX_Diabetes, Age, Sex, and CDX_Cog
summary(m1) # show results

predict(m1, df.test, type = "prob") # show the predicted odds of each subject's subtype
cm <- table(predict(m1), df.test$Subtype); cm # create table showing the predicted versus actual values for subtype

1-sum(diag(cm))/sum(cm) # calculate the % error of model

z <- summary(m1)$coefficients/summary(m1)$standard.errors # create a ratio of coefficient to standard error; necessary to get p values
p <- (1 - pnorm(abs(z), 0, 1)) * 2 # perform a 2-tailed z test to determine p values for each predictor for each outcome option, in relation to reference outcome [here, "typical"]
p # print them p's!
```
### comparing models
```{r multinom1.1}
set.seed(0001)
df.test$Subtype <- relevel(df.test$Subtype, ref = "typical")
# create models
m.base <- multinom(data = df.test, Subtype ~  Age + Sex)

m.diabetes <- multinom(data = df.test, Subtype ~ CDX_Diabetes + Age + Sex + ID_Hispanic + ID_Race_Black)

m.hypertension <- multinom(data = df.test, Subtype ~ CDX_Hypertension + Age + Sex + ID_Hispanic + ID_Race_Black)

m.BMI <- multinom(data = df.test, Subtype ~ OM_BMI + Age + Sex + ID_Hispanic + ID_Race_Black)

m.combined <- multinom(data = df.test, Subtype ~ CDX_Diabetes + CDX_Hypertension + OM_BMI + Age + Sex)

m.interction <- multinom(data = df.test, Subtype ~  CDX_Diabetes*OM_BMI + Age + Sex + ID_Hispanic + ID_Race_Black)

# comparison

models <- list(m.base, m.diabetes, m.hypertension, m.BMI, m.combined, m.interction)

model.names <- c('m.base', 'm.diabetes', 'm.hypertension', 'm.BMI', 'm.combined', 'm.interction')

aictab(cand.set = models, modnames = model.names)

```
#### Diabetes
```{r multinom1.2}
set.seed(0001)
summary(m.diabetes) 
predict(m.diabetes, df.test, type = "prob") 
cm <- table(predict(m.diabetes), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.diabetes)$coefficients/summary(m.diabetes)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### BMI
```{r multinom1.3}
set.seed(0001)
summary(m.BMI) 
predict(m.BMI, df.test, type = "prob") 
cm <- table(predict(m.BMI), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.BMI)$coefficients/summary(m.BMI)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### Hypertension
```{r multinom1.3.2}
set.seed(011)
summary(m.hypertension) 
predict(m.hypertension, df.test, type = "prob") 
cm <- table(predict(m.hypertension), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.hypertension)$coefficients/summary(m.hypertension)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### Interaction
```{r multinom1.4}
set.seed(011)
summary(m.interction) 
predict(m.interction, df.test, type = "prob") 
cm <- table(predict(m.interction), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.interction)$coefficients/summary(m.interction)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```


## limbic predom reference
```{r multinom2}
set.seed(0001)
df.test$Subtype <- relevel(df.test$Subtype, ref = "limbic predominant")
m1 <- multinom(data = df.test, Subtype ~ OM_BMI + CDX_Diabetes + Age + Sex + CDX_Cog)
summary(m1) 

predict(m1, df.test, type = "prob")
cm <- table(predict(m1), df.test$Subtype); cm 

1-sum(diag(cm))/sum(cm)

z <- summary(m1)$coefficients/summary(m1)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2; p
```
### comparing models
```{r multinom1.1.2}
# create models
set.seed(0001)
m.base2 <- multinom(data = df.test, Subtype ~  Age + Sex)

m.diabetes2 <- multinom(data = df.test, Subtype ~ CDX_Diabetes + Age + Sex + ID_Hispanic + ID_Race_Black)

m.hypertension2 <- multinom(data = df.test, Subtype ~ CDX_Hypertension + Age + Sex + ID_Hispanic + ID_Race_Black)

m.BMI2 <- multinom(data = df.test, Subtype ~ OM_BMI + Age + Sex + ID_Hispanic + ID_Race_Black)

m.combined2 <- multinom(data = df.test, Subtype ~ CDX_Diabetes + CDX_Hypertension + OM_BMI + Age + Sex)

m.interaction2 <- multinom(data = df.test, Subtype ~  Age*Sex)

# comparison

#models <- list(m.base2, m.diabetes2, m.hypertension2, m.BMI2, m.combined2, m.interaction2)

#model.names <- c('m.base', 'm.diabetes', 'm.hypertension', 'm.BMI', 'm.combined', 'm.interaction')

#aictab(cand.set = models, modnames = model.names)

```
#### Diabetes
```{r multinom2.2}
set.seed(0001)
summary(m.diabetes2) 
predict(m.diabetes2, df.test, type = "prob") 
cm <- table(predict(m.diabetes2), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.diabetes2)$coefficients/summary(m.diabetes2)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### BMI
```{r multinom2.3}
summary(m.BMI2) 
predict(m.BMI2, df.test, type = "prob") 
cm <- table(predict(m.BMI2), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.BMI2)$coefficients/summary(m.BMI2)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### Hypertension
```{r multinom2.4}
set.seed(011)
summary(m.hypertension2) 
predict(m.hypertension2, df.test, type = "prob") 
cm <- table(predict(m.hypertension2), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.hypertension2)$coefficients/summary(m.hypertension2)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```

# Histogram funk
These are not reported or used for the abstract, but demonstrate that Typical-patterned subjects have more intense tau pathology (via PET) than other groups. The difference is where it is relatively concentrated. 
```{r hist}
ggplot(df.mciad, aes(x=Hippocampus.SUVR.scaled.to.Ref, color = Subtype)) + 
  geom_histogram(bins = 100,fill = "white") + 
  labs(title = "Hippocampal tau SUVR by subtype")
  
ggplot(df.mciad, aes(x=Neocortex.SUVR.scaled.to.Ref, color = Subtype)) + 
  geom_histogram(bins = 100,fill = "white") + 
  labs(title = "Neocortical tau SUVR by subtype")
```

# Comparing AD patterns to CU
```{r compare}
# create the CU group
df.norm <- subset(df.all, df.all$CDX_Cog == "0")

# paired t tests
## examining the hippocampus
t.test(df.norm$Hippocampus.SUVR.scaled.to.Ref, df.hs$Hippocampus.SUVR.scaled.to.Ref)
t.test(df.norm$Hippocampus.SUVR.scaled.to.Ref, df.typical$Hippocampus.SUVR.scaled.to.Ref)
t.test(df.norm$Hippocampus.SUVR.scaled.to.Ref, df.lp$Hippocampus.SUVR.scaled.to.Ref)
t.test(df.norm$Hippocampus.SUVR.scaled.to.Ref, df.submed$Hippocampus.SUVR.scaled.to.Ref)

## examining the neocortex
t.test(df.norm$Neocortex.SUVR.scaled.to.Ref, df.lp$Neocortex.SUVR.scaled.to.Ref)
t.test(df.norm$Neocortex.SUVR.scaled.to.Ref, df.typical$Neocortex.SUVR.scaled.to.Ref)
t.test(df.norm$Neocortex.SUVR.scaled.to.Ref, df.typical$Neocortex.SUVR.scaled.to.Ref)
t.test(df.norm$Neocortex.SUVR.scaled.to.Ref, df.submed$Neocortex.SUVR.scaled.to.Ref)
```

# Whole Hippo volume
```{r compare2}
# analyzing whole hippo volume as measured by hippodeep
summary(df.norm$whole_hippo)
summary(df.submed$whole_hippo)
summary(df.typical$whole_hippo)
summary(df.hs$whole_hippo)
summary(df.lp$whole_hippo)

# t tests
t.test(df.norm$whole_hippo, df.submed$whole_hippo) # NS
t.test(df.norm$whole_hippo, df.typical$whole_hippo) # Significant, norm > typical -- expected!
t.test(df.norm$whole_hippo, df.lp$whole_hippo) # NS -- unexpected!
t.test(df.norm$whole_hippo, df.hs$whole_hippo) # Significant, norm > HS -- unexpected!
t.test(df.hs$whole_hippo, df.lp$whole_hippo) # NS -- unexpected!
```
