---
title: "Tau-defined subtypes by sex and metabolics"
author: "Brandon Hall"
output:
  prettydoc::html_pretty:
    theme: hpster
    highlight: github
    toc: true
    number_sections: true
---

# Loading libraries
```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2) # you know this legend
library(plyr) 
library(dplyr)
library(ggsignif) # for creating plots with sigificance added
library(nnet) # for creating multinomial logistic regression models
library(AICcmodavg) # for model comparison
library(tidyverse)
```

# Getting set up
## Loading data

This chunk loads the required CSVs and merges them into one dataframe
```{r data, echo=FALSE, message=TRUE, warning=FALSE}
setwd("~/Documents/BraskieLab/My projects/Tau subtypes x metabolics/")
df.subtypes <- read.csv("Subject tau subtype PET data v3.csv")
colnames(df.subtypes)[1] <- "Med_ID"
df.HABS <- read.csv("~/Documents/BraskieLab/HABS data/Data Release 3/HD 1  50+ Merged v2.csv")
df.all<- merge(df.subtypes, df.HABS, by = "Med_ID")
#write.csv(df.all, file = "~/Documents/BraskieLab/My projects/Tau subtypes x metabolics/subtype+HABS_merged.csv", row.names = FALSE)
```

### Change variable types

This chunk changes the variable types as needed for subsequent analyses
```{r var_types, echo=FALSE, message=FALSE, warning=FALSE}
df.all$CDX_Cog <- as.factor(df.all$CDX_Cog)
df.all$Age <- as.numeric(df.all$Age)
df.all$ID_Gender <- as.factor(df.all$ID_Gender)
df.all$CDX_Diabetes <- as.factor(df.all$CDX_Diabetes)
df.all$Diabetes <- as.factor(df.all$Diabetes)
df.all$CDX_Hypertension <- as.factor(df.all$CDX_Hypertension)
df.all$ID_Hispanic <- as.factor(df.all$ID_Hispanic)
df.all$ID_Race_Black <- as.factor(df.all$ID_Race_Black)
df.all$APOE4_Positivity <- as.factor(df.all$APOE4_Positivity)
df.all$simoa_plasma_Total_Tau <- as.numeric(df.all$simoa_plasma_Total_Tau)
df.all$Final_subtype <- factor(df.all$Final_subtype)
df.all$X01_AB_FBB_pos <- as.factor(df.all$X01_AB_FBB_pos)
df.all$APOE4_Positivity <- as.factor(df.all$APOE4_Positivity)
df.all$OM_BMI <- as.double(df.all$OM_BMI)
df.all$Subtype <- factor(df.all$Subtype, ordered = FALSE)
df.all$Age_log <- log(df.all$Age)
df.all$OM_BMI_log <- log(df.all$OM_BMI)
df.all$X01_L_hippocampus <- as.numeric(df.all$X01_L_hippocampus)
df.all$X01_R_hippocampus <- as.numeric(df.all$X01_R_hippocampus)
df.all$whole_hippo <- (df.all$X01_L_hippocampus + df.all$X01_R_hippocampus)/2
```

### Subset from MCI & AD

This chunk create subset dataframes from the prior, unified dataframe
```{r subsetting2}
# only MCI and AD subjects
df.mciad <- subset(df.all,df.all$CDX_Cog != "0")
df.mciad$OM_BMI <- as.double(df.mciad$OM_BMI)
df.mciad$CDX_Cog <- factor(df.mciad$CDX_Cog, ordered = FALSE)
df.females <- subset(df.mciad, df.mciad$ID_Gender == "1")
df.males <- subset(df.mciad, df.mciad$ID_Gender == "0")
df.submed <- subset(df.mciad, df.mciad$Final_subtype == "0")
df.typical <- subset(df.mciad, df.mciad$Final_subtype == "1")
df.lp <- subset(df.mciad, df.mciad$Final_subtype == "2")
df.hs <- subset(df.mciad, df.mciad$Final_subtype == "3")
df.MA <- subset(df.mciad, df.mciad$ID_Hispanic != "1")
df.NHW <- subset(df.mciad, df.mciad$ID_Hispanic == "1") 
df.black <- subset(df.mciad, df.mciad$ID_Race_Black == "1")
df.typical_f <- subset(df.typical, df.typical$ID_Gender=="1") # create typical female-only subset
df.typical_m <- subset(df.typical, df.typical$ID_Gender=="0") # create typical male-only subset
df.lp_f <- subset(df.lp, df.lp$ID_Gender=="1") # create lp female-only subset
df.lp_m <- subset(df.lp, df.lp$ID_Gender=="0") # create lp male-only subset
df.hs_f <- subset(df.hs, df.hs$ID_Gender=="1") # create hs female-only subset
df.hs_m <- subset(df.hs, df.hs$ID_Gender=="0") # create hs male-only subset
df.submed_f <- subset(df.submed, df.submed$ID_Gender=="1") # create submed female-only subset
df.submed_m <- subset(df.submed, df.submed$ID_Gender=="0") # create submed male-only subset

#write.csv(df.mciad, file = "~/Documents/BraskieLab/My projects/Tau subtypes x metabolics/MCI_AD.csv", row.names = FALSE)

df.test <- subset(df.mciad, select = c("Med_ID", "Subtype", "OM_BMI", "OM_BMI_log", "ID_Hispanic", "ID_Race_Black", "CDX_Diabetes", "Diabetes", "CDX_Hypertension", "Age", "Sex", "CDX_Cog", "Ratio.of.Hippocampal.SUVR.to.Neocortical.SUVR", "ID_Hispanic", "ID_Race_Black")) # create a subsetted data frame containing only what variables I need

df.test <- df.test[complete.cases(df.test) ,] # removed any subjects who have incomplete data
```

# Barcharts

## 1: Sex

This chunk provides the barcharts for visualizing the difference in sex between subtypes
```{r sex_plot}
# subtype x sex
ggplot(df.test, aes(x = Subtype, fill = Sex)) + 
  geom_bar(width = 0.4, position = "dodge") + 
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Sex by tau subtype", fill = "Sex", x = "Subtype") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
## 2: Diabetes

This chunk provides the barcharts for visualizing the difference in diabetic diagnoses between subtypes
```{r sex_diab_plot}

# subtype x diabetes
ggplot(df.test, aes(x = Subtype, fill = Diabetes)) + 
  geom_bar(width = 0.6, position = "dodge") + 
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Diabetic diagnosis by subtype", fill = "Diabetes", x = "Subtype") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

##3: hypertension

```{r hypertenseplot}

ggplot(df.mciad, aes(x = Subtype, fill = CDX_Hypertension)) + geom_bar(width = 0.4, position = "dodge") + labs(title = "hypertension diagnosis by subtype", fill = "hypertension diagnosis", x = "subtype") + ylim(0,100)
```
##4: ID_Black

```{r blackplot}

ggplot(df.test, aes(x = Subtype, fill = ID_Race_Black)) + 
  geom_bar(width = 0.4, position = "dodge") + 
  labs(title = "Black identity by subtype", fill = "Self-report identity: black", x = "subtype") + 
  ylim(0,100) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

##5: ID_Hispanic

```{r hispplot}

ggplot(df.test, aes(x = Subtype, fill = ID_Hispanic)) + 
  geom_bar(width = 0.4, position = "dodge") + 
  labs(title = "Hispanic identity by subtype", fill = "Self-report identity: Hispanic", x = "subtype") + 
  ylim(0,100) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
# Chi square test

``` {r chisq}
set.seed(1001)
chisq.test(df.mciad$ID_Hispanic, df.mciad$Subtype, correct = FALSE, p = rep(1/length(df.mciad$Subtype), length(df.mciad$Subtype)), simulate.p.value = TRUE, B = 1000)

chisq.test(df.mciad$ID_Race_Black, df.mciad$Subtype, correct = FALSE, p = rep(1/length(df.mciad$Subtype), length(df.mciad$Subtype)), simulate.p.value = TRUE, B = 1000)
```

# multinomial logistic regression

All models include covariates of age and cognitive status

## Typical reference
### comparing models
```{r multinom1.1}
set.seed(0001)
df.test$Subtype <- relevel(df.test$Subtype, ref = "typical")
# create models
m.diabetes <- multinom(data = df.test, Subtype ~ CDX_Diabetes + Age + Sex + ID_Hispanic + ID_Race_Black)

m.hypertension <- multinom(data = df.test, Subtype ~ CDX_Hypertension + Age + Sex + ID_Hispanic + ID_Race_Black)

m.BMI <- multinom(data = df.test, Subtype ~ OM_BMI + Age + Sex + ID_Hispanic + ID_Race_Black)

m.interction <- multinom(data = df.test, Subtype ~  CDX_Diabetes*OM_BMI + Age + Sex + ID_Hispanic + ID_Race_Black)
```
#### Diabetes
```{r multinom1.2}
set.seed(0001)
summary(m.diabetes) 
predict(m.diabetes, df.test, type = "prob") 
cm <- table(predict(m.diabetes), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.diabetes)$coefficients/summary(m.diabetes)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### BMI
```{r multinom1.3}
set.seed(0001)
summary(m.BMI) 
predict(m.BMI, df.test, type = "prob") 
cm <- table(predict(m.BMI), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.BMI)$coefficients/summary(m.BMI)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### Hypertension
```{r multinom1.3.2}
set.seed(011)
summary(m.hypertension) 
predict(m.hypertension, df.test, type = "prob") 
cm <- table(predict(m.hypertension), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.hypertension)$coefficients/summary(m.hypertension)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### Interaction

Add for per the feedback of Dr. Yaffe. No sig interaction. 
```{r multinom1.4}
set.seed(011)
summary(m.interction) 
predict(m.interction, df.test, type = "prob") 
cm <- table(predict(m.interction), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.interction)$coefficients/summary(m.interction)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```


## limbic predom reference
```{r multinom2}
set.seed(0001)
df.test$Subtype <- relevel(df.test$Subtype, ref = "limbic predominant")
```
### comparing models
```{r multinom1.1.2}
# create models
set.seed(0001)

m.diabetes2 <- multinom(data = df.test, Subtype ~ CDX_Diabetes + Age + Sex + ID_Hispanic + ID_Race_Black)

m.hypertension2 <- multinom(data = df.test, Subtype ~ CDX_Hypertension + Age + Sex + ID_Hispanic + ID_Race_Black)

m.BMI2 <- multinom(data = df.test, Subtype ~ OM_BMI + Age + Sex + ID_Hispanic + ID_Race_Black)
```
#### Diabetes
```{r multinom2.2}
set.seed(0001)
summary(m.diabetes2) 
predict(m.diabetes2, df.test, type = "prob") 
cm <- table(predict(m.diabetes2), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.diabetes2)$coefficients/summary(m.diabetes2)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### BMI
```{r multinom2.3}
summary(m.BMI2) 
predict(m.BMI2, df.test, type = "prob") 
cm <- table(predict(m.BMI2), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.BMI2)$coefficients/summary(m.BMI2)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```
#### Hypertension
```{r multinom2.4}
set.seed(011)
summary(m.hypertension2) 
predict(m.hypertension2, df.test, type = "prob") 
cm <- table(predict(m.hypertension2), df.test$Subtype); cm
1-sum(diag(cm))/sum(cm)
z <- summary(m.hypertension2)$coefficients/summary(m.hypertension2)$standard.errors 
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```

# T-tests
## Age
```{r t.test}
t.test(df.males$Age, df.females$Age) # total group t-mciad -- NS
t.test(df.typical$Age, df.submed$Age) # typical vs sub-med -- significant!
t.test(df.typical$Age, df.hs$Age) # typical vs hippocampal sparing -- NS
t.test(df.typical$Age, df.lp$Age) # typical vs limbic predominant -- NS
t.test(df.hs$Age, df.lp$Age) # hippocampal sparing vs limbic predomiant -- significant!

```
## BMI
```{r t.test2}
t.test(df.hs$OM_BMI, df.typical$OM_BMI)
t.test(df.lp$OM_BMI, df.typical$OM_BMI)
t.test(df.hs$OM_BMI, df.lp$OM_BMI)
t.test(df.submed$OM_BMI, df.typical$OM_BMI)
```
## Race/Ethnicity
```{r t.test3}
df.typical$ID_Race_Black <- as.numeric(df.typical$ID_Race_Black)
df.hs$ID_Race_Black <- as.numeric(df.hs$ID_Race_Black)
df.lp$ID_Race_Black <- as.numeric(df.lp$ID_Race_Black)
t.test(df.hs$ID_Race_Black, df.typical$ID_Race_Black)
t.test(df.hs$ID_Race_Black, df.lp$ID_Race_Black)
t.test(df.lp$ID_Race_Black, df.typical$ID_Race_Black)

df.typical$ID_Hispanic <- as.numeric(df.typical$ID_Hispanic) 
df.hs$ID_Hispanic <- as.numeric(df.hs$ID_Hispanic)
df.lp$ID_Hispanic <- as.numeric(df.lp$ID_Hispanic)
t.test(df.hs$ID_Hispanic, df.typical$ID_Hispanic)
t.test(df.hs$ID_Hispanic, df.lp$ID_Hispanic)
t.test(df.lp$ID_Hispanic, df.typical$ID_Hispanic)

```
